---
title: 'IS 604 – Homework #1'
author: "J. Hamski"
date: "September 3, 2016"
output: pdf_document
---

```{r}
library(ggplot2)
```

1.1) Name several entities, attributes, activities, events, and state variables for the following systems:

a. Cafeteria

*entities*: diners, cash register attendants, cooks  
*attributes*: tables vacancies, line wait time  
*activities*: sitting down, eating, waiting to check out  
*events*: arrive at seat, depart seat, checking out     
*state variables*: time, diner number    


b. Grocery store  
*entities*: shelves, items on shelves (UPC codes)       
*attributes*: quantity of product available    
*activities*: restocking, picking out items     
*events*: taking an item off a shelf    
*state variables*: time, shift     

c. laudromat  
*entities*: customers, washers, dryers     
*attributes*: wash/dry time   
*activities*: washing, drying   
*events*: wash/dry started, wash/dry ended   
*state variables*: time, wash/dry cycles  

d. Fast-food restaurant
*entities*: cars in a drive through window       
*attributes*: order preparation time   
*activities*: ordering, preparing order     
*events*: order entered, order prepared     
*state variables*: car number, time   

e. Hospital emergency room
*entities*: patients   
*attributes*: time of treatment, severity (priority for treatment)   
*activities*: treatment, waiting      
*events*: entering into queue,  triage, wait time, treatment start, treatment end     
*state variables*: time, waiting order, priority  

f. Taxicab company with 10 taxis
*entities*: cars, riders, drivers     
*attributes*: hours of operation, breaks     
*activities*: picking up fares, driving them to destination     
*events*: pickup, drop-off     
*state variables*: taxi occupancy, time, trip length  

g. Automobile assembly line
*entities*: cars, workers, robots       
*attributes*: percent assembled, maintenance requirements, shift changes  
*activities*: assembling a single car       
*events*: assembly of a specific section of car, fixing a specific part     
*state variables*: time, begin assebly, end essembly, ready for assembly action   

2.1)

```{r}
set.seed(243)
# time unit is minutes

between.arrivals <- c(0,60,120,180)
between.arrivals.prob <- c(0.23,0.37,0.28,0.12)


shop.model <- function(max.customers){
  
  customer <- 1
  arrival.time <- 0
  service.time <- 50
  service.time.begins <- 25
  service.time.ends <- NULL
  time.in.system <- 25
  time.in.queue <- NULL
  int.arrival.time <- NULL
  
  for (i in 1:(max.customers-1)) {
    
    
    customer <- c(customer, customer[i] + 1)
    #print(customer)
    
    sim.service.time <- round(rnorm(n = 1, mean = (50), sd = (8)),0)
    service.time <- c(service.time, sim.service.time)
    
    int.arrival.time.customer <- sample(x = between.arrivals, size = 1, prob = between.arrivals.prob)
    int.arrival.time <- c(int.arrival.time, int.arrival.time.customer)
    
    arrival.time <- c(arrival.time, arrival.time[i-1] + int.arrival.time.customer)
    
    service.time.begins <- c(service.time.begins, max(arrival.time[i], service.time.ends[i - 1]))
    
    service.time.ends <- c(service.time.ends, service.time.begins[i] + service.time[i])
    
    time.in.system <- c(time.in.system, service.time.ends[i] - arrival.time[i])

    time.in.queue <- c(time.in.queue, arrival.time[i] - service.time.begins[i])
    
    #print(service.time[i])
  }
  output <- list(customer, time.in.queue, service.time, time.in.system)
  return(output)
}

a <- shop.model(10)
```

Average time in queue:
```{r}
mean(a[[1]])
```

Average processing time:
```{r}
mean(a[[2]])
```

Maximum time in system:
```{r}
max(a[[3]])
```

This model shows reasonable results for the average and max processing times, but certainly has an error (that I failed to debug) since time in system shows occasonal negative numbers. 

2.2)

```{r}
#bagel units are dozens

customers.day <- c(8,10,12,14)
customer.prob <- c(0.35,0.30,0.25,0.10)

dozens.ordered <- c(1,2,3,4)
ordered.prob <- c(0.4,0.3,0.2,0.1)

bagel.demand.sim <- function(quantity, sim.days){
  
  day <- 1
  demand <- NULL
  revenue <- NULL
  lost.profit <- NULL
  salvage <- NULL
  daily.cost <- quantity * 5.8
  daily.profit <- NULL

  
  for (i in 1:(sim.days)) {
    day <- c(day, day[i] + 1)
    #print(day)
    #print(i)
    
    #start demand subroutine
    customers <- sample(x = customers.day, prob = customer.prob, size = 1)
    bagels.sold.day <- NULL
    for (j in 1:(customers)) {
      bagels.sold.day <- c(bagels.sold.day, sample(x = dozens.ordered, prob = ordered.prob, size = 1))
    }
    bagels.sold <- sum(bagels.sold.day)
    #end demand subroutine
    
    #print(bagels.sold)
    
    demand <- c(demand, bagels.sold)
    
    revenue <- c(revenue, bagels.sold * 8.4)

    lost.profit.day <- ifelse(quantity < bagels.sold, (bagels.sold - quantity) * 2.6, 0)
    lost.profit <- c(lost.profit, lost.profit.day)
    
    salvage.day <- ifelse(quantity > bagels.sold, (quantity - bagels.sold)*4.2, 0)
    salvage <- c(salvage, salvage.day)
    
    daily.profit <- c(daily.profit, bagels.sold * 8.4 - lost.profit.day + salvage.day - daily.cost)
    
  }
  return(mean(daily.profit))
}
```

```{r}
max <- 5
min <- 60

quantities <- seq(max, min, 5)
```

```{r}
five.day.sim.results <- sapply(quantities, FUN=bagel.demand.sim, sim.days = 5)
plot(five.day.sim.results)
```


Limiting to five days simulation only makes sense with Excel. Here, 1,000 days are simulated for each 5 dozen bin. 
```{r}
onethousand.day.sim.results <- sapply(quantities, FUN=bagel.demand.sim, sim.days = 1000)
plot(onethousand.day.sim.results)
```


2.4)

```{r}
interarrival <- c(15,20,25,30,35)
inter.prob <- c(0.14,0.22,0.43,0.17,0.04)

service.time <- c(5,15,25,35,45)
service.prob <- c(0.12,0.35,0.43,0.06,0.04)

one.taxi <- function(){
  
  
}

```



2.5) 

```{r}
rand.var.sim <- function(){
  X <- rnorm(1, mean = 100, sd = 100)
  Y <- rnorm(1, mean = 300, sd = 225)
  Z <- rnorm(1, mean = 40, sd = 64)
  
  W <- (X + Y)/Z
  return(W)  
}
```

```{r}
rand.var.results <- replicate(50, rand.var.sim())
qplot(rand.var.results, geom = "histogram", binwidth = 3) 
```

Note that 50 is not really enough replications to show the full variance of this distribution. Simulating the random variable 1000 times shows this random variable has a tendency to produce some extreme values.   

```{r}
rand.var.results <- replicate(1000, rand.var.sim())
qplot(rand.var.results, geom = "histogram", binwidth = 3) 
var(rand.var.results)
```


2.6)

```{r}
B.values <- c(0,1,2,3,4)
B.probs <- rep(1/5, length(B.values))

C.values <- c(10,20,30,40)
C.probs <- c(0.1,0.25,0.5,0.15)

D.sim <- function() {
  A <- rnorm(1, mean = 100, sd = 400)
  B <- sample(x = B.values, prob = B.probs, size = 1)
  C <- sample(x = C.values, prob = C.probs, size = 1)
  
  D <- (A - 25*B) / (2*C)
  return(D)
}
```

```{r}
D.sim.results <- replicate(100, D.sim())

mean(D.sim.results)
```

2.7

```{r}
# output goal: average number of lost sales per week

inventory.sim <- function(sim.weeks){
  day <- 1
  inventory <- 18
  
  first.day.demand <- rnorm(1, mean = 5, sd = 1.5)
  lost.sales <- ifelse(first.day.demand > inventory, first.day.demand - inventory, 0)
  
  
  
  #simulation time unit is 1 day. Loop starts on day 2
  for (i in 1:(sim.weeks*7)) {
    
  }

}

a <- inventory.sim(5)
a
```

2.8)

Here I try a different approach than previous problems. I first generate the required set of probability distribution defined random variables, then loop through to perform the loading calculations. 
```{r}
#time unit is minutes


elevator.sim <- function(sim.min) {
  A <- runif(sim.min, min = 3, max = 7)
  B <- rep(6, sim.min)
  C <- sample(x = c(2, 3), prob = c(0.33, 0.67), replace = TRUE, size = sim.min)
  
  time <- 0
  
  
  #calculate loading
  while (time < 60) {
    time <- time + 1
    print(time)
    
    
    
  }
  
  
}

elevator.sim(60)


```


