---
title: 'IS 604 – Homework #1'
author: "J. Hamski"
date: "September 3, 2016"
output: pdf_document
---

1.1) Name several entities, attributes, activities, events, and state variables for the following systems:

a. Cafeteria

*entities*: diners  
*attributes*: tables vacancies  
*activities*: sitting down, eating  
*events*: arrive at seat, depart seat  
*state variables*: number of diners  


b. Grocery store  
*entities*: shelves, items on shelves (UPC codes)     
*attributes*:  
*activities*:   
*events*:   
*state variables*:   

c. laudromat
*entities*:   
*attributes*: 
*activities*:   
*events*:   
*state variables*:

d. Fast-food restaurant
*entities*:   
*attributes*: 
*activities*:   
*events*:   
*state variables*:

e. Hospital emergency room
*entities*: patients   
*attributes*: severity of ailment 
*activities*: treatment   
*events*:   
*state variables*:

f. Taxicab company with 10 taxis
*entities*:   
*attributes*: 
*activities*:   
*events*:   
*state variables*:

g. Automobile assembly line
*entities*: cars     
*attributes*: percent assembled
*activities*: assembling a single car       
*events*: assembly of a specific section of car, fixing a specific part     
*state variables*: 

2.1)

```{r}
# time unit is minutes

between.arrivals <- c(0,60,120,180)
between.arrivals.prob <- c(0.23,0.37,0.28,0.12)


shop.model <- function(max.customers){
  
  customer <- 1
  arrival.time <- 0
  service.time <- 25
  service.time.begins <- 
  service.time.ends <- NULL
  time.in.system <- 0
  time.in.queue <- 0
  int.arrival.time <- 0
  
  for (i in 1:(max.customers - 1)) {
    customer <- c(customer, customer[i]+1)
    #print(customer)
    
    sim.service.time <- round(rnorm(n = 1, mean = (50), sd = (8)),0)
    service.time <- c(service.time, sim.service.time)
    
    int.arrival.time <- c(int.arrival.time, sample(x = between.arrivals, size = 1, prob = between.arrivals.prob))
    
    arrival.time <- c(arrival.time, arrival.time[i] + int.arrival.time[i])
    
    service.time.begins <- c(service.time.begins, max(arrival.time[i], service.time.ends[i - 1]))
    
    service.time.ends <- c(service.time.ends, service.time.begins[i] + service.time[i])
    
    time.in.system <- c(time.in.system, arrival.time[i] - service.time.ends[i])

    time.in.queue <- c(time.in.queue, service.time.begins[i] - arrival.time[i])
    
    #print(service.time)
  }
  output <- list(time.in.queue, service.time, time.in.system)
  return(output)
}

a <- shop.model(10)

```

2.2)

```{r}
customers.day <- c(8,10,12,14)
customer.prob <- c(0.35,0.30,0.25,0.10)

dozens.ordered <- c(1,2,3,4)
ordered.prob <- c(0.4,0.3,0.2,0.1)

bagel.sim <- function(){
  day = 1
  
  while(day <=5){
    customers <- sample(x = customers.day, size = 1, prob = customer.prob)
    
    bagels.ordered = NULL
    
    bagels.ordered = for(i in 1:customers){i * sample(x = dozens.ordered, size = 1, prob = ordered.prob)}
    
    day = day + 1
    
  }
}

bagel.sim()
```


2.4)

```{r}
interarrival <- c(15,20,25,30,35)
inter.prob <- c(0.14,0.22,0.43,0.17,0.04)

service.time <- c(5,15,25,35,45)
service.prob <- c(0.12,0.35,0.43,0.06,0.04)

one.taxi <- function(){
  
  
}

```



2.5) 

2.7)

2.8)

